---
title: "Naive Bayes from scratch"
author: "Joe D"
date: "9/30/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Naive Bayes Algorithm in R

This is an example implementation of the naive bayes learning algorithm in R.  The code originated from <https://gist.github.com/dansondergaard/7062952>.

The algorithm is designed after the wikipedia entry for naive bayesand uses the data set from that example.

```{r}

# naivebayes.txt
#
# sex height weight  footsize
# male  6    180	12
# male	5.92 190	11
# male	5.58	170	12
# male	5.92	165	10
# female 5	100	6
# female 5.5	150	8
# female 5.42	130	7
# female 5.75	150	9

library("data.table")
library("e1071") # for titanic data used below
library(plyr)

train <- function (dt) {
  return(dt[,list(meanHeight=mean(height), varHeight=sd(height),
                  meanWeight=mean(weight), varWeight=sd(weight),
                  meanFootsize=mean(footsize), varFootsize=sd(footsize)), by=sex])
}

classify <- function (classifier, sample) {
  posterior <- function (sample, class_prior, class) {
    p_height <- dnorm(sample$height, class$meanHeight, class$varHeight)
    p_weight <- dnorm(sample$weight, class$meanWeight, class$varWeight)
    p_footsize <- dnorm(sample$footsize, class$meanFootsize, class$varFootsize)
    return(class_prior * p_height * p_weight * p_footsize)
  }
  
  class_male <- classifier[which(classifier$sex == 'male'),]
  class_female <- classifier[which(classifier$sex == 'female'),]
  
  prior_male <- 0.5
  prior_female <- 0.5
  
  return(list(male=posterior(sample, prior_male, class_male),
              female=posterior(sample, prior_female, class_female)))
}
```

We run the algorithm by loading a dataset and executing 'train' as below.  This learns the classification procedure.  Then we can apply data using 'classify' and evaluate the model produced by 'train'.

```{r}
training_set <- data.table(read.table('naivebayes.txt', header=TRUE))
sample = data.table(height=6, weight=130, footsize=8)

classifier <- train(training_set)
result = classify(classifier, sample)

cat('posterior(male) =', result$male)
cat('\nposterior(female) =', result$female)
```
We can rerun this algorithm using the Titanic dataset as described below.  This example is derived from <https://www.r-bloggers.com/understanding-naive-bayes-classifier-using-r/>
```{r}
data(Titanic)
#Save into a data frame and view it
Titanic_df=as.data.frame(Titanic)
Titanic_df
```
In this dataset, we have a summary of all the combinations of the above variables from a subset of the full Titanic set.  We need to expand this to use it for training:

```{r}
#Creating data from table
repeating_sequence=rep.int(seq_len(nrow(Titanic_df)), Titanic_df$Freq) #This will repeat each combination equal to the frequency of each combination
#Create the dataset by row repetition created
Titanic_dataset=data.table(Titanic_df[repeating_sequence,])
#We no longer need the frequency, drop the feature
Titanic_dataset$Freq=NULL

head(Titanic_dataset)
```
Now we can try to write code to fit this model.

```{r}
train_titanic <- function(dt) {
  tblAge <- table(dt$Survived, dt$Age)
#  print(tblAge)
  tblSex <- table(dt$Survived, dt$Sex)
#  print(tblSex)
  tblClass = table(dt$Survived, dt$Class)
#  print(tblClass)
#  names(tblClass) <- tblAge
#  names(tblSex) <- tblAge
#  return (dt[, cbind(tblAge, tblSex, tblClass)])
  return (list(age=tblAge, sex=tblSex, class=tblClass))
}

```

```{r}
predict_Titanic <- function (classifier, sample) {
  posterior <- function (sample, class_prior, class) {
    p_age <- class$age/rowSums(class$age)
    p_sex <- class$sex/rowSums(class$sex)
    p_class <- class$class/rowSums(class$class)
    print(class_prior)
    return(class_prior * p_age * p_sex * p_class)
  }
  
#  class_survived <- list(fAge = classifier$age[2,],fSex=classifier$sex[2,], fClass=classifier$class[2,])
#  class_died <- list(fAge = classifier$age[1,],fSex=classifier$sex[1,], fClass=classifier$class[1,])
  
  prior_survived <- 0.5
  prior_died <- 0.5
  
  result <- posterior(sample, rbind(prior_survived, prior_died), classifier)
  return(result)
#  return(list(survived=posterior(sample, prior_survived, class_survived),
#              died=posterior(sample, prior_died, class_died)))
}
```

```{r}
sample_titanic = data.table(Age="Adult", Sex="Male", Class="Crew")

classifier <- train_titanic(Titanic_dataset)
classifier
result = predict_Titanic(classifier, sample)

#cat('posterior(survived) =', result$survived)
#cat('\nposterior(died) =', result$died)
```